---
title: "Assignment1"
author: "Daniel Yowell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Total questions completed: 15/15

# 1. Summarize how I will workout your final grade for the course. Give percentages etc. Give my grading scale also e.g. What percentage is an A etc

MATH 4753 Grade Visualization:  

• 30% comes from the final.  
• 20% comes from midterms (2 worth 10% each).  
• 15% comes from a cumulative 4 assignments.  
• 10% comes from projects (a third of this grade comes from Project 1, the rest from Project 2).  
• 10% comes from labs.  
• 10% comes from in-class quizzes.  
• 5% comes from online Canvas quizzes on each chapter.  

Some points may also be derived from minilab reports and additional class exercises.  

Letter grades are assigned as:  

• A: 90s  
• B: 80s  
• C: 60s-70s  
• D: 50s  
• F: < 50  

There are no curves.

```{r}
grades <- c(
  "Final" = 30, 
  "Midterms" = 20, 
  "Assignments" = 15, 
  "Projects" = 10, 
  "Labs" = 10, 
  "In-class quizzes" = 10, 
  "Online quizzes" = 5
  )

# Customizing the pie chart
pie(grades, labels = names(grades), col = rainbow(length(grades)), main = "Daniel Yowell - MATH 4753 Grade Distribution")
legend("topright", cex=0.8, legend = names(grades), fill = rainbow(length(grades)))

```

# 2. A biologist wants to make a coplot of LENGTH Vs WEIGHT given RIVER*SPECIES for fish caught in the Tennessee river and recorded in the DDT.csv data set, so that each point is colored according to the variable MILE which is treated as a factor (Qualitative variable).

### (a) Make the coplot as the biologist required Hint: Use coplot(), Lab 1, the code provided, and plotting options pch and col to differentiate the MILE variable. You should be able to produce something like what is shown below

```{r}
ddt <- read.csv("DDT.csv")

m=with(ddt, as.numeric(factor(MILE))) # A 
print(ddt$MILE)
print(factor(ddt$MILE))
print(m)
length(unique(m)) #B

coplot(LENGTH ~ WEIGHT | RIVER*SPECIES, data = ddt, pch = 1, col = m)
```

### (b) Interpret the lower left three conditional plots.

All three coplots display the relationship between WEIGHT and LENGTH of CCATFISH.  
Each coplot only contains the CCATFISH populations from the rivers FCM, LCM, and SCM (reading from left to right, respectively.)  
The different colors of the points represent the MILE variable for each CCATFISH. However, the value does not vary between members of the same RIVER, so the colors remain constant within each plot (unlike the TRM fish, which vary widely).

### (c) What does line A do?

Line A swaps each value with its ranking in a list from smallest to largest (for instance, 1 is left unchanged, 3 becomes 2, 5 becomes 3, 275 becomes 4, etc.).  

To do this, it uses the factor() function, which converts the vector into a factor object. Next, as_numeric() replaces each value with its respective factor level. 

### (d) What does line B do?

Line B counts the number of unique factor levels (in this case, 17).

### (e) Why are the top six plots empty?

The data does not contain any SMBUFFALO and LMBASS fish outside of the TRM river. These six plots would otherwise display that, but here they are instead left blank.

### (f) What is the mean value of DDT found in the sample of CCATFISH caught in the FCM river?

```{r}
library(dplyr)
subset <- ddt %>% filter(SPECIES == 'CCATFISH' & RIVER == 'FCM')
mean(subset$DDT)
```

# 3. MS 1.14 - pg 8

### National Bridge Inventory. All highway bridges in the United States are inspected periodically for structural deficiency by the Federal Highway Administration (FHWA). Data from the FHWA inspections are compiled into the National Bridge Inventory (NBI). Several of the nearly 100 variables maintained by the NBI are listed below. Classify each variable as quantitative or qualitative.

a. Length of maximum span (feet): QUANTITATIVE
b. Number of vehicle lanes: QUANTITATIVE
c. Toll bridge (yes or no): QUALITATIVE
d. Average daily traffic: QUANTITATIVE
e. Condition of deck (good, fair, or poor): QUALITATIVE
f. Bypass or detour length (miles): QUANTITATIVE
g. Route type (interstate, U.S., state, county, or city): QUALITATIVE

# 4. MS page 12,13 Read pages 12 and 13 about random sampling designs and answer the following:

### (a) What are the names of the four random sampling designs (1 simple and 3 more complex).
### (b) Give a brief description of each.

# 5. MS 1.15 - pg 15 – Use sample(...,replace=FALSE), if mtbe is the dataframe then we need a random sample of the rows. If v is a vector containing a random sample of row indices then mtbe[v,] will be the random sample.

# 12. MS 2.73 - pg 70

# 13. MS 2.80 - pg 72

### Mongolian desert ants. The Journal of Biogeography (Dec. 2003) published an article on the first comprehensive study of ants in Mongolia (Central Asia). Botanists placed seed baits at 11 study sites and observed the ant species attracted to each site. Some of the data recorded at each study site are provided in the table at the top of p. 73.

### a. Find the mean, median, and mode for the number of ant species discovered at the 11 sites. Interpret each of these values.

### b. Which measure of central tendency would you recommend to describe the center of the number of ant species distribution? Explain.

### c. Find the mean, median, and mode for the total plant cover percentage at the 5 Dry Steppe sites only.

### d. Find the mean, median, and mode for the total plant cover percentage at the 6 Gobi Desert sites only.

### e. Based on the results, parts c and d, does the center of the total plant cover percentage distribution appear to be different at the two regions?

```{r}
#Mode function
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

# 14. MS 2.84 - pg 74



# 15. Using the ddt data set re-create the plot below using ggplot. Make sure your plot is titled with your name. NB – You MUST use ggplot()

```{r}
library(ggplot2)
ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) + geom_boxplot() + ggtitle("Daniel Yowell's Q15 Boxplot")
```








