---
title: "testing1"
author: "Daniel Yowell"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is a testing document.
# It contains experiments, explanations of R commands, and answers to previous quizzes.

# Cool R Facts

Did you know? You can pass a data frame into the curly brackets to save some time.
(skips read.csv() function)
```{r ddt}
getwd()
summary(ddt)
```

## Task 2.5
## to be deleted

```{r ddt}
column <- ddt$SPECIES
unique_count <- length(unique(column))
unique_count
```

# QUIZ 1: SQUARE BRACKETEERING

1. In the ddt data frame find the entry in the first row and second column. 
```{r ddt}
head(ddt)
```

2. Find the amount of DDT in the flesh of the fish which has the longest length. Hint which.max() might help
```{r ddt}
ddt[which.max(ddt$LENGTH),"DDT"]
```

3. How many fish have a length > 49 and < 50? Hint: Use an "&"
```{r ddt}
ddt[ddt$LENGTH > 49  & ddt$LENGTH < 50,]
```

4. How many fish have a length > 49 and weight < 1200?
```{r ddt}
ddt[ddt$LENGTH > 49  & ddt$WEIGHT < 1200,]
```

# QUIZ 2: WRANGLED UP IN KNOTS

1. Find the mean WEIGHT of SMBUFFALO fish to 4 decimal places
```{r ddt}
ddt %>% filter(SPECIES == "SMBUFFALO") %>% summarize(n = mean(WEIGHT))
```

2. Find the mean LENGTH of CCATFISH which have DDT > 33 to 2 decimal places
```{r ddt}
ddt %>% filter(DDT > 33 & SPECIES == "CCATFISH") %>% summarize(n = mean(LENGTH))
```

3. Find the standard deviation of DDT for fish that are either  LMBASS  or SMBUFFALO to 4 decimal places
```{r ddt}
# METHOD 1: Wrangler
ddt %>% filter(SPECIES == "LMBASS" | SPECIES == "SMBUFFALO" ) %>% summarize(sd_DDT = sd(DDT))
# METHOD 2: Classic
sd(with(ddt, ddt[SPECIES == "LMBASS" | SPECIES == "SMBUFFALO",]$DDT))
```

4. Find the number of fish that have a LENGTH > 40  and WEIGHT > 1000
```{r ddt}
# Method 1 - WRANGLER APPROACH
# 1. pass ddt into filter function
# 2. in filter function, sort by LENGTH > 40 and WEIGHT > 1000
# 3. finally, summarize(). the n() function counts the number of rows, 
#    assigning it to a new column n.
# 4. optionally, you can print n to the console with the code `n`
ddt %>% filter(LENGTH > 40 & WEIGHT > 1000) %>% summarize(n = n())
# Method 2 - SQUARE BRACKET APPROACH
# 1. we want the output of a length() function
# 2. with() is kinda clunky, but basically the first parameter specifies
#    the intended data frame so that we don't have to recite it later
# 3. ddt[...] searches ddt for the given parameters
# 4. '1' selects the first column of data (any column works since we just
#    care to recognize a given row's existence)
# - Passing the with() string into length() gives us the size of the vector
length(with(ddt, ddt[LENGTH > 40 & WEIGHT > 1000, 1]))
```

# QUIZ 3:

1. Find the mean LENGTH to 4 decimal places.
```{r ddt}
mean(ddt$LENGTH)
```

2. Find the standard deviation of WEIGHT to 4 decimal places.
```{r ddt}
sd(ddt$WEIGHT)
```

How many fish are longer than 50?
```{r ddt}
ddt %>% filter(LENGTH > 50) %>% summarize(n = n())
```

How many SMBUFFALO fish are longer than 50?
```{r ddt}
ddt %>% filter(LENGTH > 50 & SPECIES == "SMBUFFALO") %>% summarize(n = n())
```
```{r ddt}
ddt %>% filter(LENGTH > 50 & SPECIES == "SMBUFFALO")
```
# QUIZ 3

z_i is the number of standard deviations that x_i is away from the sample mean.

[TRUE]

Using z, find the value of the WEIGHT of fish that is defined as an outlier

```{r ddt}
x <- ddt$WEIGHT
z=(x-mean(x))/sd(x)
ddt$WEIGHT[abs(z)>3]
```


Using the same data set (DDT.csv) find all values of WEIGHT that are possible outliers.

Of these possible outliers write down the smallest!

```{r ddt}
ddt$WEIGHT[abs(z)>=2 & abs(z)<=3]
```
