---
title: "Lab2"
author: "Daniel Yowell"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

Get the working directory

```{r}
getwd()
```

# Task 2

Read EPAGAS into data frame

```{r}
data <- read.csv("EPAGAS.csv")
head(data)
```

# Task 3

Make the object mpg, the number of miles per gallon vector.

```{r}
mpg <- data$MPG
```

If $z_i=\frac{(x_i-\bar{x})}{s_x}$, then $\bar{z} = 0$ and $s_z^2 = 1$. 
Transform the mpg variable to z and verify these results.

```{r}
# Subtract mean from mpg, then divide by standard deviation
z=(mpg-mean(mpg))/sd(mpg)
print(z)
# The mean of z is extremely close to zero
print(mean(z))
# The square of the standard deviation of z is 1
print(sd(z)^2)
```

Using z, find the values of mpg that are possible outliers.

```{r}
z[abs(z)>=2 & abs(z)<=3]
```

Using z, find the values of mpg that are defined as outliers.

```{r}
z[abs(z)>3]
```

Using the lattice package, construct a dotplot.
(red = outlier, blue = possible outlier)

```{r}
# outliers = red, possible outliers = blue, rest = black
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))  

library(lattice)
dp <- dotplot(mpg,col=mycol, main="Daniel Yowell's MPG Dotplot") 
print(dp)
```

# Task 4

Make a boxplot of the mpg variable.

```{r}
boxplot(data$MPG, main="Daniel Yowell's MPG Boxplot", ylab="MPG", horizontal=TRUE, col="black", notch = TRUE)

# Median value
median_val <- median(data$MPG)

# Add the median line
abline(h = median_val)
```

Q: Using Chebyshev’s theorem predict the proportion of data within 2 standard deviations of the mean of the mpg data.

  A: Chebyshev's theorem says that at least 75% of the data lies within 2 standard deviations of the mean.

Use R to calculate the exact proportion within 2 standard deviation of the mean.


```{r}
mean_mpg <- mean(mpg)
print(mean_mpg)
sd2 <- sd(mpg) * 2
print(sd2)

upperbound <- mean_mpg + sd2
lowerbound <- mean_mpg - sd2

subset <- mpg[mpg > lowerbound  & mpg < upperbound]
print(length(subset) / length(mpg))
```

Q: Does Chebyshev agree with the data?

  A: We have shown that 96% of the data lies within 2 standard deviations of the mean. Thus, Chebyshev agrees with the data.

Q: Now use the Empirical rule. What proportion (according to the rule) of the data should be within 2 standard deviations of the mean?

  A: The Empirical rule says that 95% of the data should be within 2 standard deviations of the mean.

Q: How well does it correspond?

  A: It's a close call, but the 96% we found earlier is indeed greater than 95%. Thus, the Empirical rule holds up.

Q: Is the Empirical rule valid in this case? Why?

  A: It is acceptable to use the Empirical rule here, since our data forms the shape of a single mound (as shown in the histogram below).

```{r}
hist(mpg, 
     main = "Daniel Yowell's MPG Histogram", 
     xlab = "mpg", 
     ylab = "frequency")
```
