---
title: "my-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(FALL224753yowe0004)
```

This my mycltu, which returns a histogram of the central limit for a given uniform distribution.

```{r}
FALL224753yowe0004::mycltu(n=20,iter=100000)
```

This is ntickets, the function we worked on for Project 1. Here is a demonstration of the scenario with 200 total seats, a 0.02 chance of overbooking, and a 0.95 probability for each passenger showing:

```{r}
FALL224753yowe0004::ntickets(200, 0.02, 0.95)
```

This is the maximum likelihood function we created in Lab 10. We still need to manually create the log function before passing it as a parameter, but otherwise it works exactly as intended.

```{r}
logbin=function(x,param) log(dbinom(x,prob=param,size=10))
FALL224753yowe0004::mymaxlik(x = c(1, 2, 3, 4, 5, 6, 7, 8),
                             param = seq(0,1,length=1000),
                             lfun = logbin,
                             xlab = expression(pi),
                             main = "Binomial",
                             cex.main = 1
                             )
```

This is how to create a confidence interval using myci(). To demonstrate, we'll use the fire data set included with our package.

```{r}
fire <- FALL224753yowe0004::fire
FALL224753yowe0004::myci(d = fire$DISTANCE, alpha = 0.05)
```
